---
title: "Claims Frequency Modeling — End-to-end Pipeline (GLM/ZIP/GBM + Interpretation)"
output:
  html_notebook:
    toc: true
    toc_depth: 3
params:
  project_root: ""
  run_pipeline: true
  seed: 1
  k_folds: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

# ----------------------------
# Project root finder
# ----------------------------
find_project_root <- function(start_dir = getwd(), max_up = 10) {
  cur <- normalizePath(start_dir, winslash = "/", mustWork = FALSE)
  for (i in 0:max_up) {
    if (file.exists(file.path(cur, "scripts", "01_setup.R"))) return(cur)
    parent <- normalizePath(file.path(cur, ".."), winslash = "/", mustWork = FALSE)
    if (identical(parent, cur)) break
    cur <- parent
  }
  return(NA_character_)
}

# ----------------------------
# Decide root dir (works from /notebooks or root)
# ----------------------------
# 1) If param exists, use it
PROJECT_ROOT <- NULL
if (exists("params") && !is.null(params$project_root) && nzchar(params$project_root)) {
  PROJECT_ROOT <- normalizePath(params$project_root, winslash = "/", mustWork = FALSE)
}

# 2) Otherwise, detect it automatically (starting from notebook location)
if (is.null(PROJECT_ROOT) || !nzchar(PROJECT_ROOT) || is.na(PROJECT_ROOT)) {
  PROJECT_ROOT <- find_project_root(getwd(), max_up = 10)
}

if (is.na(PROJECT_ROOT)) {
  stop(
    "Project root not found.\n",
    "Open this notebook from inside the project folder, or set params$project_root.\n",
    "Expected to find: scripts/01_setup.R"
  )
}

# IMPORTANT: set both knit root.dir and working dir
knitr::opts_knit$set(root.dir = PROJECT_ROOT)
setwd(PROJECT_ROOT)

cat("Project root:", PROJECT_ROOT, "\n")

# ----------------------------
# Load pipeline config
# ----------------------------
source(file.path("scripts", "01_setup.R"))

# Optional: override seed / kfolds from params if you want
if (exists("params") && !is.null(params$seed) && !is.na(params$seed)) {
  SEED <- as.integer(params$seed)
}
if (exists("params") && !is.null(params$k_folds) && !is.na(params$k_folds)) {
  K_FOLDS <- as.integer(params$k_folds)
}

set.seed(SEED)
cat("Notebook SEED:", SEED, "\n")
cat("Notebook K_FOLDS:", K_FOLDS, "\n")
```

## What this notebook does

This notebook is a **reproducible wrapper** around your existing scripts (01--08).\
It will:

1.  Run the full pipeline (import → cleaning → feature engineering → models → interpretation).
2.  Load the saved outputs (tables/models/plots) produced by your scripts.
3.  Show a quick "what to look at" summary so you have something presentable.

------------------------------------------------------------------------

## 1) Run pipeline (01 → 08)

```{r run-pipeline}
RUN_PIPELINE <- isTRUE(params$run_pipeline)

if (RUN_PIPELINE) {
  cat("Running pipeline scripts 02–08...\n")

  # 02 - import & clean
  source(file.path("scripts", "02_import_clean.R"))

  # 03 - feature engineering
  source(file.path("scripts", "03_feature_engineering.R"))

  # 04 - glm models (Poisson / NB / ZIP)
  source(file.path("scripts", "04_glm_models.R"))

  # 05 - gbm models
  source(file.path("scripts", "05_gbm_models.R"))

  # 06 - interpretation (SHAP etc.)
  source(file.path("scripts", "06_interpretation_SHAP.R"))

  # 07 - ALE/ICE/PDP 
  source(file.path("scripts", "07_gbm_effect_plots.R"))
  
  # 08 - ICE
  source(file.path("scripts", "08_interpretation_LIME.R"))

  cat("Pipeline finished.\n")
} else {
  cat("RUN_PIPELINE = FALSE. Skipping execution; will only load outputs if they exist.\n")
}
```

------------------------------------------------------------------------

## 2) Load processed data (train/test)

```{r load-data}
train_path <- file.path(DIR_DATA_PROCESSED, "train_data.rds")
test_path  <- file.path(DIR_DATA_PROCESSED, "test_data.rds")

if (file.exists(train_path) && file.exists(test_path)) {
  train_data <- readRDS(train_path)
  test_data  <- readRDS(test_path)

  cat("Train rows:", nrow(train_data), "\n")
  cat("Test rows :", nrow(test_data), "\n")
} else {
  cat("Processed train/test not found yet.\n",
      "Expected:\n - ", train_path, "\n - ", test_path, "\n", sep = "")
}
```

------------------------------------------------------------------------

## 3) Model comparison tables

### GLM / ZIP comparison (CV + Holdout)

```{r glm-tables}
glm_cv_path <- file.path(DIR_OUTPUT_TABLES, "glm_cv_results.csv")
glm_holdout_path <- file.path(DIR_OUTPUT_TABLES, "glm_holdout_results.csv")

if (file.exists(glm_cv_path)) {
  glm_cv <- read.csv(glm_cv_path)
  glm_cv
} else {
  cat("Missing:", glm_cv_path, "\n")
}

if (file.exists(glm_holdout_path)) {
  glm_holdout <- read.csv(glm_holdout_path)
  glm_holdout
} else {
  cat("Missing:", glm_holdout_path, "\n")
}
```

### GBM results (grid + holdout)

```{r gbm-tables}
gbm_grid_path <- file.path(DIR_OUTPUT_TABLES, "gbm_grid_results.csv")
gbm_holdout_path <- file.path(DIR_OUTPUT_TABLES, "gbm_holdout_results.csv")
gbm_varimp_path <- file.path(DIR_OUTPUT_TABLES, "gbm_variable_importance.csv")

if (file.exists(gbm_holdout_path)) {
  gbm_holdout <- read.csv(gbm_holdout_path)
  gbm_holdout
} else {
  cat("Missing:", gbm_holdout_path, "\n")
}

if (file.exists(gbm_grid_path)) {
  gbm_grid <- read.csv(gbm_grid_path)
  head(gbm_grid)
} else {
  cat("Missing:", gbm_grid_path, "\n")
}

if (file.exists(gbm_varimp_path)) {
  gbm_varimp <- read.csv(gbm_varimp_path)
  head(gbm_varimp)
} else {
  cat("Missing:", gbm_varimp_path, "\n")
}
```

------------------------------------------------------------------------

## 4) Interpretation outputs (SHAP / Break-down / LIME / PDP-ALE-ICE)

```{r list-outputs}
list_files_safe <- function(path, pattern = NULL) {
  if (!dir.exists(path)) return(character(0))
  list.files(path, pattern = pattern, full.names = TRUE)
}

cat("Tables folder:\n")
print(list_files_safe(DIR_OUTPUT_TABLES))

cat("\nModels folder:\n")
print(list_files_safe(DIR_OUTPUT_MODELS))

cat("\nPlots folder:\n")
if (exists("DIR_OUTPUT_PLOTS")) {
  print(list_files_safe(DIR_OUTPUT_PLOTS))
} else {
  # fallback if you use a different constant name
  print(list_files_safe(file.path("output", "plots")))
}
```

------------------------------------------------------------------------

## Session info

```{r session-info}
sessionInfo()
```
